A linear programming approach to the tracking of partials

### 2017-04-25 paper/paper.tex
Added conclusion and did one read-through.

\begin{abstract}
    A new approach to the tracking of sinusoidal chirps using linear programming
    is proposed. It is demonstrated that the classical algorithm of
    \cite{mcaulay1986speech} is greedy and exhibits exponential complexity for
    long searches. A linear programming (LP) formulation to find the best $L$ paths
    in a lattice is described and its complexity is analysed. Finally it is
    demonstrated that the new LP formulation outperforms the classical algorithm
    in the tracking of sinusoidal chirps in high levels of noise --- both in
    computational complexity and robustness.
\end{abstract}

\begin{keywords}
partial tracking, linear programming, optimization, additive synthesis
\end{keywords}

@@

The latter techniques seeking globally optimal sets of paths incur great
computational cost due to the large number of possible solutions. For this reason, in
this paper we propose a linear programming \textit{relaxation}-formulation of
the optimal path-set problem based on
an algorithm for the tracking of multiple objects in video
\cite{jiang2007linear}. It will be shown that this algorithm has favourable
asymptotic complexity and performs well on the tracking of chirps in high levels
of noise.

@@

In this section, we present the McAulay-Quatieri method of peak matching. It is
conceptually simple and a set of short paths can be computed quickly, but
it can be sensitive to spurious peaks and is optimal only in the sense that
the set of paths computed contains the best path possible --- the quality of the
other paths may be compromised under this criterion.

In \cite[p.~748]{mcaulay1986speech} the peak matching algorithm is described in
a number of steps; we summarize them here in a way comparable with the linear
programming formulation to be presented shortly. In that paper, the parameters
of each data point are the instantaneous amplitude, phase, and frequency but we
will allow for arbitrary parameter sets $\theta$.  Define a distance function
$\mathcal{D} \left( \theta_{i},\theta_{j} \right)$ that computes the
similarity between $2$ sets of parameters. We will now consider a method that
finds $L$ tuples of parameters that are closest.

@@

All nodes in $\BS{x}^{\ast}$ will have at most one incoming connection or
otherwise no connections, a constraint that can be enforced through the
following linear inequality. Define $\BS{A}^{\text{I}} \in
\mathbb{R}^{R_{\text{I}} \times P}$ with $R_{\text{I}} = \sum_{k=1}^{K-1}
N_{k}$, the number of nodes in all the frames excluding the first. We sum all
the connections into the node $r_{\text{I}} + N_{0}$ represented by the
respective entry in $\BS{x}$ through an inner product with the $r_{\text{I}}$th
row in $\BS{A}^{\text{I}}$ and require that this sum be between $0$ and $1$,
i.e.,
\begin{equation}
    \BS{A}^{\text{I}}_{r_{\text{I}},\mathcal{B}(p)} = \begin{cases}
        1 & \text{if } p_{1} = r_{\text{I}}+N_{0} \\
        0 & \text{otherwise}
    \end{cases}, 0 \leq r_{\text{I}} < R_{\text{I}}, p \in \rho
\end{equation}
and
\begin{equation}
    \BS{0} \leq \BS{A}^{\text{I}}\BS{x} \leq \BS{1}
\end{equation}
Similarly, to constrain the number of outgoing connections into each node, we
define $R_{\text{O}} = \sum_{k=0}^{K-2} N_{k}$ and
$\BS{A}^{\text{O}} \in \mathbb{R}^{R_{\text{O}} \times P}$ with
\begin{equation}
    \BS{A}^{\text{O}}_{r_{\text{O}},\mathcal{B}(p)} = \begin{cases}
        1 & \text{if } p_{0} = r_{\text{O}} \\
        0 & \text{otherwise}
    \end{cases}, 0 \leq r_{\text{O}} < R_{\text{O}}, p \in \rho
\end{equation}
and
\begin{equation}
    \BS{0} \leq \BS{A}^{\text{O}}\BS{x} \leq \BS{1}
\end{equation}

To ensure no breaks in the paths it is required that the number of incoming
connections into a given node equal the number of outgoing connections for the
$R_{\text{B}} = \sum_{k=1}^{K-2} N_{k}$ nodes potentially having both incoming
and outgoing connections.
\begin{equation}
    \BS{A}^{\text{B}}_{r_{\text{B}}} = \BS{A}^{\text{B}}_{r_{\text{B}}} -
    \BS{A}^{\text{B}}_{r_{\text{B}}+N_{0}} \text{ for rows } 0 \leq r_{\text{B}}
    < R_{\text{B}}
\end{equation}
and
\begin{equation}
    \label{eq:cxnbalcon}
    \BS{A}^{\text{B}}\BS{x} = \BS{0}
\end{equation}

@@

The complete LP to find the $L$ best disjoint paths through a lattice described
by node connections $\rho$ is then
\[
    \min_{\BS{x}} \cLP^{T} \BS{x} 
\]
subject to
\[
    \BS{G}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{I}} \\
        \BS{A}^{\text{O}} \\
        -\BS{I}
    \end{bmatrix} \BS{x} \leq
    \begin{bmatrix}
        \BS{1} \\
        \BS{1} \\
        \BS{0}
    \end{bmatrix}
\]
\begin{equation}
    \label{eq:lpprogfull}
    \BS{A}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{B}} \\
        \BS{A}^{\text{C}}_{K-2}
    \end{bmatrix} \BS{x} = 
    \begin{bmatrix}
        \BS{0} \\
        L
    \end{bmatrix}
\end{equation}
where $\BS{I}$ is the identity matrix.

@@

Although the matrices involved in (\ref{eq:lpprogfull}) are large, only a small
fraction of their values are non-zero. As stated above, the number of variables in $\BS{x}$ is
$P = (K-1)N^{2}$. Even though matrices $\BS{A}^{\text{I}}, \BS{A}^{\text{O}} \in
\mathbb{R}^{N(K-1) \times P}$, each contains only $P$ non-zero
entries.  $\BS{A}^{\text{B}} \in \mathbb{R}^{N(K-1) \times P}$ but contains only
$2N^{2}(K-2)$ non-zero entries while $\BS{A}^{\text{C}}_{K-2} \in \mathbb{R}^{P}$
contains merely $N$. The $\BS{x} \geq \BS{0}$ constraint requires a matrix
with $P$ non-zero entries. The total memory complexity is therefore $4P +
2N^{2}(K-2)$ floating-point numbers.

@@

We search for partial tracks using both the greedy and LP strategies. Both
algorithms use the distance metric $\mathcal{D}_{\text{pr.}}$ between two parameters sets:
\[
    \mathcal{D}_{\text{pr.}} \left( \theta_{i}^{k},
    \theta_{j}^{k+1} \right) = \left( \omega_{i}^{k} +
    \psi_{i}^{k} H - \omega_{j}^{k+1} \right)
\]
which is the error in predicting $j$th frequency in frame $k+1$ from the $i$th
parameters in frame $k$. For the greedy method, the search for partial paths is
restricted to two frames ahead, i.e., paths of length $K_{\text{MQ}}=3$ are sought, otherwise the
computation becomes intractable. For the LP
method the search is carried out over all frames ($K_{\text{LP}}=28$). %TODO: Is it?
The cost thresholding values are ${\Delta_{\text{MQ}} = \Delta_{\text{LP}} =
0.1}$. For both methods, the search is restricted
to nodes between frequencies 250 to 2000 Hz.

Figure~\ref{plot:mq_lp_compare_chirps}
shows discovered partial trajectories for signals at various SNRs. It is seen
that while the greedy method starts performing poorly at an SNR of -6db, the LP method
still gives plausible partial trajectories. The LP method returns paths spanning
all $K$ frames, due to the constraints. The McAulay-Quatieri method in general does not,
but longer paths can be formed in a straightforward way after the initial short path
search step \cite{mcaulay1986speech}.

@@

In this paper we reformulated the classical greedy algorithm of McAulay and
Quatieri and showed that it can be seen as a greedy algorithm for finding the $L$
shortest paths in a lattice. An algorithm was then proposed minimizing the sum
of the $L$ paths, using a linear programming approach. The complexity of the new
algorithm was shown to be generally less than the Viterbi-based algorithms and
the McAulay-Quatieri algorithm for large $K$.
It was shown on synthetic signals that the new approach finds plausible paths in
lattices with a large number of spurious nodes.

@@

The proposed algorithm, while asymptotically faster than other partial tracking
algorithms, is still not fast. In many situations where computational resources
are limited, a McAulay-Quatieri method search over many sets of small $K$ works
sufficiently well. However in high amounts of noise the algorithm proposed here
is robust while still of tractable complexity.

%TODO: Improvements and extensions

>>>>

\begin{abstract}
    A new approach to the tracking of sinusoidal chirps using linear programming
    is proposed. It is demonstrated that the classical algorithm of
    \cite{mcaulay1986speech} is greedy and exhibits exponential complexity for
    long searches, while approaches based on the Viterbi algorithm exhibit
    factorial complexity \cite{depalle1993tracking} \cite{wolf1989finding}. A
    linear programming (LP) formulation to find the best $L$ paths in a lattice
    is described and its complexity is shown to be less than previous
    approaches. Finally it is demonstrated that the new LP formulation
    outperforms the classical algorithm in the tracking of sinusoidal chirps in
    high levels of noise.
\end{abstract}

\begin{keywords}
partial tracking, linear programming, optimization, additive synthesis, atomic
decomposition, regularized approximation
\end{keywords}

@@

The latter techniques seeking globally optimal sets of paths incur great
computational cost due to the large number of possible solutions. For this reason, in
this paper we propose a linear programming \textit{relaxation} formulation of
the optimal path-set problem based on
an algorithm for the tracking of multiple objects in video
\cite{jiang2007linear}. It will be shown that this algorithm has favourable
asymptotic complexity and performs well on the tracking of chirps in high levels
of noise.

@@

In this section, we present the McAulay-Quatieri method of peak matching. It is
conceptually simple and a set of short paths can be computed quickly, but it can
be sensitive to spurious peaks and its complexity becomes unwieldly for long
searches.

In \cite[p.~748]{mcaulay1986speech} the peak matching algorithm is described in
a number of steps; we summarize them here in a way comparable with the linear
programming formulation to be presented shortly. In that paper, the parameters
of each data point are the instantaneous amplitude, phase, and frequency but
here we allow for arbitrary parameter sets $\theta$.  Define a distance function
$\mathcal{D} \left( \theta_{i},\theta_{j} \right)$ that computes the similarity
between $2$ sets of parameters. We will now consider a method that finds $L$
tuples of parameters that are closest.

@@

All nodes in $\BS{x}^{\ast}$ will have at most one incoming connection or
otherwise no connections, a constraint that can be enforced through the
following linear inequality: define $\BS{A}^{\text{I}} \in
\mathbb{R}^{R_{\text{I}} \times P}$ with $R_{\text{I}} = \sum_{k=1}^{K-1}
N_{k}$, the number of nodes in all the frames excluding the first. We sum all
the connections into the node $r_{\text{I}} + N_{0}$ represented by the
respective entry in $\BS{x}$ through an inner product with the $r_{\text{I}}$th
row in $\BS{A}^{\text{I}}$ and require that this sum be between $0$ and $1$,
i.e.,
\begin{equation}
    \BS{A}^{\text{I}}_{r_{\text{I}},\mathcal{B}(p)} = \begin{cases}
        1 & \text{if } p_{1} = r_{\text{I}}+N_{0} \\
        0 & \text{otherwise}
    \end{cases}, 0 \leq r_{\text{I}} < R_{\text{I}}, p \in \rho
\end{equation}
and
\begin{equation}
    \BS{0} \leq \BS{A}^{\text{I}}\BS{x} \leq \BS{1}
\end{equation}
Similarly, to constrain the number of outgoing connections into each node, we
define $R_{\text{O}} = \sum_{k=0}^{K-2} N_{k}$ and
$\BS{A}^{\text{O}} \in \mathbb{R}^{R_{\text{O}} \times P}$ with
\begin{equation}
    \BS{A}^{\text{O}}_{r_{\text{O}},\mathcal{B}(p)} = \begin{cases}
        1 & \text{if } p_{0} = r_{\text{O}} \\
        0 & \text{otherwise}
    \end{cases}, 0 \leq r_{\text{O}} < R_{\text{O}}, p \in \rho
\end{equation}
and
\begin{equation}
    \BS{0} \leq \BS{A}^{\text{O}}\BS{x} \leq \BS{1}
\end{equation}

To forbid breaks in the paths it is required that the number of incoming
connections into a given node equal the number of outgoing connections for the
$R_{\text{B}} = \sum_{k=1}^{K-2} N_{k}$ nodes potentially having both incoming
and outgoing connections.
\begin{equation}
    \BS{A}^{\text{B}}_{r_{\text{B}}} = \BS{A}^{\text{B}}_{r_{\text{B}}} -
    \BS{A}^{\text{B}}_{r_{\text{B}}+N_{0}} \text{ for rows } 0 \leq r_{\text{B}}
    < R_{\text{B}}
\end{equation}
and
\begin{equation}
    \label{eq:cxnbalcon}
    \BS{A}^{\text{B}}\BS{x} = \BS{0}
\end{equation}

@@

The complete LP to find the $L$ best disjoint paths through a lattice described
by node connections $\rho$ is then
\[
    \min_{\BS{x}} \cLP^{T} \BS{x} 
\]
subject to
\[
    \BS{G}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{I}} \\
        \BS{A}^{\text{O}} \\
        -\BS{I}
    \end{bmatrix} \BS{x} \leq
    \begin{bmatrix}
        \BS{1} \\
        \BS{1} \\
        \BS{0}
    \end{bmatrix}
\]
\begin{equation}
    \label{eq:lpprogfull}
    \BS{A}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{B}} \\
        \BS{A}^{\text{C}}_{K-2}
    \end{bmatrix} \BS{x} = 
    \begin{bmatrix}
        \BS{0} \\
        L
    \end{bmatrix}
\end{equation}
where $\BS{I}$ is the identity matrix. A proof that the solution
$\boldsymbol{x}^{\ast}$ will have entries equal to either $0$ or $1$ can be
found in \cite[p.~167]{parker1988discrete}.

@@

Although the matrices involved in (\ref{eq:lpprogfull}) are large, only a small
fraction of their values are non-zero. Matrices $\BS{A}^{\text{I}},
\BS{A}^{\text{O}} \in \mathbb{R}^{N(K-1) \times P}$, but each contains only $P$
non-zero entries.  Furthermore $\BS{A}^{\text{B}} \in \mathbb{R}^{N(K-1) \times
P}$ but contains only $2N^{2}(K-2)$ non-zero entries while
$\BS{A}^{\text{C}}_{K-2} \in \mathbb{R}^{P}$ contains merely $N$. The $\BS{x}
\geq \BS{0}$ constraint requires a matrix with $P$ non-zero entries. The total
memory complexity including the entries in $\cLP$ and the right-hand-sides of
(\ref{eq:lpprogfull}) is $2N^{2}(K-2) + 4P + 2N(K-1) + N + 1$ non-zero
floating-point numbers.

@@

We search for partial tracks using both the greedy and LP strategies. Both
algorithms use the distance metric $\mathcal{D}_{\text{pr.}}$ between two parameters sets:
\begin{equation}
    \label{eq:examplecostfun}
    \mathcal{D}_{\text{pr.}} \left( \theta_{i}^{k},
    \theta_{j}^{k+1} \right) = \left( \omega_{i}^{k} +
    \psi_{i}^{k} H - \omega_{j}^{k+1} \right)
\end{equation}
which is the error in predicting $j$th frequency in frame $k+1$ from the $i$th
parameters in frame $k$. For the greedy method, the search for partial paths is
restricted to two frames ahead, i.e., paths of length $K_{\text{MQ}}=3$ are sought, otherwise the
computation becomes intractable. For the LP
method the search is carried out over all frames ($K_{\text{LP}}=28$).
The cost thresholding values are ${\Delta_{\text{MQ}} = \Delta_{\text{LP}} =
0.1}$. For both methods, the search is restricted
to nodes between frequencies 250 to 2000 Hz.

Figure~\ref{plot:mq_lp_compare_chirps}
shows discovered partial trajectories for signals at various SNRs. It is seen
that while the greedy method starts performing poorly at an SNR of -6 dB, the LP method
still gives plausible partial trajectories. The LP method returns paths spanning
all $K$ frames, due to the constraints. The McAulay-Quatieri method in general does not,
but longer paths can be formed in a straightforward way after the initial short path
search step \cite{mcaulay1986speech}.

@@

In this paper we reformulated the classical partial tracking technique of McAulay and
Quatieri and showed that it can be seen as a greedy algorithm for finding the $L$
shortest paths in a lattice. An algorithm was then proposed minimizing the sum
of the $L$ paths, using a linear programming approach. The complexity of the new
algorithm was shown to be generally less than the Viterbi-based methods and
the McAulay-Quatieri algorithm for large $K$.
It was shown on synthetic signals that the new approach finds plausible paths in
lattices with a large number of spurious nodes.

@@

In its current form, the path search may choose undesirable paths if a
convenient node is missing from the following frame. An extension could consider
nodes some number of frames ahead.

The proposed algorithm, while asymptotically faster than other partial tracking
algorithms, is still not fast. In situations where computational resources
are limited, a McAulay-Quatieri method search over many sets of small $K$ works
sufficiently well. However in high amounts of noise the algorithm proposed here
is robust while still of tractable complexity.

It may be possible to improve the performance of the algorithm by the use of
different cost functions and regularization. The cost function
(\ref{eq:examplecostfun}) could be extended to encourage similarity between
frequency slopes or amplitude information. Each metric should be scaled
according to its desired contribution to the cost.

There may also be a way to extract individual paths through the use of auxiliary
variables in (\ref{eq:lpprogfull}). If so, path specific costs such as overall
smoothness or fit to a particular model could be incorporated.

In any case, it would be interesting to further investigate programming
relaxations encouraging underlying discrete structures plausible for audio
in the framework of regularized approximation. These structures are closer to
ground-truth structures for speech (text) and music (the musical score).
